Storing messages with multipartite neural cliques

### 2014-07-23 Articles/article.tex
...

	We study the properties of the network after storing messages in it. We are interested in how much edges have been added to the network and if the process introduced confusion between messages.

@@

	We first compute the theoretical density of the network, that is to mean the probability for a random edge $(i, i')$ to be present in the network (i.e. $W_{ii'} = 1$).

@@

	Multipartite cliques saturate quicker.

@@

@@

	\section{Maximum theoretical efficiency}
		It was shown in \cite{Palm1980} that maximal asymptotic capacity of Willshaw networks is $\log 2$. We adapt this result to multipartite clique clustered networks.

@@

	We are interested in distinct problems of messages retrieval. We detail what those problems are. Since those problems vary in hardness, we propose distinct measures for efficiencies. Efficiency measures the ratio of information effectively retrieved over the information used for storage.

@@

	A message $m$ can be recovered from a partial probe $\tilde{m}$ where some symbols, say $c_e$, with indexes $\{j_1, \dots , j_{c_e} \}$ have been replaced by the blank symbol $\intercal$, with $\intercal \not\in\mathcal{A}$ (recall $\mathcal{A}$ is the alphabet we are working on).  We have $\tilde{m} \in \left( \mathcal{A} \cup \{ \intercal \}\right)^c$.

	We can extend the mapping $f$ to $\left( \mathcal{A} \cup \{ \intercal \}\right)^c$, by associating to the symbol $\intercal$ a cluster where none of the units are activated. If we consider the binary representation $\mu$ of $m$, erased symbols corresponds to erased clusters. An erased cluster is a cluster where none of the units are activated. Formally the partial probe $\tilde{\mu}$ is such that for all $j\in [c]$ and $k \in [\ell ]$, \[ 
		\tilde{\mu}[(j-1)\ell+k] = \bigg \{ \begin{array}{cl} 0 & \mbox{if } j\in \{j_1, \dots , j_{c_e} \}\\
													\mu[(j-1)\ell+k] & \mbox{otherwise}	
	\end{array}\]

	An erasure is tantamount to lack of information, but associative memories can also retrieve messages from corrupted probes.

@@

	We propose one model of message corruption. As all symbols in $\mathcal{A}$ all contains exactly $a$ ones, symbols which do not belong to $\mathcal{A}$ can be easily detected. We therefore restrict ourselves to random replacements by symbols in $\mathcal{A}$.

	A message $m$ can be recovered from a corrupt probe $\tilde{m}$ where some symbols, say $c_e$, with indexes $\{j_1, \dots , j_{c_e} \}$ have been replaced by random elements of $\mathcal{A}$. We have $\tilde{m} \in \mathcal{A}^c$.

@@

@@

@@

	We provide a formula for error rate in case of messages with corrupted clusters.

	Messages can be corrupted, our models for errors is : given $c_e$ the number of erroneous clusters, we draw at random $c_e$ clusters and in each of these draw at random, uniformly, a new letter in the alphabet that will replace the old one (it could be the same).

	Knowing the density of the network, which follows formula (\ref{formula_density}), we can compute distributions for scores achieved by different type of units in different clusters. We can divide units in 5 groups with 3 characteristics. Subscript $g$ will indicate a unit being part of the message (it should be activated), subscript $b$ the contrary (the unit should not be activated). Subscript $a$ means the unit is already activated. Subscript $c$ corresponds to a correct cluster, $e$ to an erroneous cluster.
	Correct units in uncorrupted clusters achieve at least a score of $a(c-c_e - 1) + \gamma$. 
	Being connected to a given activated unit in a corrupted cluster occurs with probability $d$, since errors are identically distributed. Moreover corrects units see their scores shift, as do all activated units respectively thanks to being part of the original message and the memory effect $\gamma$. We can express those probability laws as : 
	For example an \textbf{a}ctivated \textbf{g}ood unit in a \textbf{c}orrect cluster starts with a base score of $ a(c - c_e - 1) + \gamma$. The probability it has to be connected to a random activated unit in an erroneous cluster is $d$. Thereby its law is :

@@

	Correct units in corrupted clusters achieve at least a score of $a(c - c_e)$. 
	$P(n_{ge} = a(c - c_e) + x) = {a (c_e - 1) \choose x} d^x (1-d)^{a (c_e - 1)-x}$

@@

	%$P(n_{be} = x) = P(n_{bc} = x) = {c - 1 \choose x} d^x (1-d)^{c- 1 -x}$
	Finally $P(n_{be} = x) = P(n_{bc} = x) = {a(c - 1) \choose x} d^x (1-d)^{a(c- 1) -x}$

@@

	The probability for all good units of a correct cluster (resp. of an erroneous cluster) to achieve a score equal or greater than $x$, with at least one having a score of $x$ is $P_{ac} (x) = \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = x)^k P(n_{agc} > x)^{a-k}$ (resp. $P_e(x) = \sum_{k = 1}^a { a \choose k } P(n_{ge} = x)^k P(n_{ge} > x)^{a-k}$)

%	\begin{align*}
%	&P_{retrieve} = \left [ \sum_{n = 1}^{a(c-1) + \gamma} \left [ \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = n)^k P(n_{agc} > n)^{a-k} \right ] P(n_{bc} < n)^{l-a} \right ]^{c - c_e}\\ 
%	 &\times \left [ \sum_{n = 1}^{a(c-1) +\gamma} \left[ \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k} \right ] P(n_{be} < n)^{l-2a} P(n_{abe} < n)^a \right ]^{c_e} \numberthis \label{from_err_th}
%	\end{align*}
		\begin{align*}
	&P_{retrieve} = \left [ \sum_{x = 1}^{a(c-1) + \gamma} P_{ac}(x) P(n_{bc} < x)^{\ell-a} \right ]^{c - c_e}\\ 
	 &\times \left [ \sum_{x = 1}^{a(c-1) +\gamma} P_e(x) P(n_{be} < x)^{\ell-2a} P(n_{abe} < x)^a \right ]^{c_e} \numberthis \label{from_err_th}
	\end{align*}

@@

	As information on a message is shared across edges with multiple activated units per cluster, such networks are more resilient to damages (for one message information is carried by $a^2 {c \choose 2}$ edges. $v_c$ will indicate a correct unit that should be activated, $v$ the other type of units.

>>>>

	We study the properties of the network once some messages have been stored in it. We are interested in how much edges have been added to the network and if the process introduced confusion between messages.

@@

	We first compute the theoretical density of the network, that is to mean the probability for an edge $(i, i')$ to be present in the network (i.e. $W_{ii'} = 1$).

@@

	Multipartite cliques saturate quicker, which was expected since they take much more edges.

@@

	\subsection{Rate of the network}

	The rate $R$ of the network measures how much information is stored in the network compared to what it takes to store it. Note it doesn't take into account how effectively messages are retrieved.

		The matrix $W$ used to store the network is binary, symmetric, with null blocks of size $\ell^2$ on the diagonal. It can therefore be stored in memory as $\frac{c(c-1) \ell^2}{2}$ bits.

		Since messages are not stored in an ordered fashion, the entropy $H(\mathcal{M})$ of the set of messages is $\log_2(\frac{({\ell \choose a}^c)^M}{M!}) \mathop{=}_{M \rightarrow +\infty} M(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2}) + O(\log_2(M)) $ (thanks to  Stirling's formula).% The ratio of information effectively stored in the network is $P_{retrieve} \times  \frac{2M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2}$.

		$R$ can be written $\frac{2H(\mathcal{M})}{c(c-1) \ell^2}$.

		A very good approximation is thereby $R = \frac{2M \left(c \log_2{\ell \choose a } - \log_2(M) + \frac{1}{\log 2} \right)}{c(c-1)\ell^2}$.

		The efficiency of the network measures how much information can be retrieved compared to how much could be at the maximum. The efficiency is called $\eta$. $\eta = P_{retrieve} R  h^{-1}$.

@@

	\subsection{Maximum theoretical efficiency}
		It was shown in \cite{Palm1980} that maximal asymptotic capacity of Willshaw networks is $\log 2$. We adapt this result to multipartite clique clustered networks.

		Asymptotic efficiency of multipartite clique networks is also $\log 2$.

@@

	We are interested in distinct problems of messages retrieval. We detail what those problems are. Since those problems vary in hardness, we propose distinct measures for efficiencies. We first compute the rate of the network. Efficiency measures the ratio of information effectively retrieved over the hardness of the problem. It can be expressed in function of the rate $R$, the retrieval rate $P_{retrieve}$ and some measure of hardness of the problem $h$.

	To account for hardness of problems, we consider an "optimal" associative memory. We are interested in retrieving messages with a few iterations at most. However an "optimal" memory could store $\mathcal{M}$ with $H(\mathcal{M})$ bits and recover messages by maximum likelihood (with ambiguities causing errors in our study, to simplify the problem). We detail the performance of such a memory for each problem, which gives use a measure for the hardness of each problem.

@@

	\subsubsection{Model}

	A message $m$ can be recovered from a partial probe $\tilde{m}$ where some symbols, say $c_e$, of indexes $\{j_1, \dots , j_{c_e} \}$, have been replaced by the blank symbol $\intercal$, with $\intercal \not\in\mathcal{A}$ (recall $\mathcal{A}$ is the alphabet we are working on).  We have $\tilde{m} \in \left( \mathcal{A} \cup \{ \intercal \}\right)^c$.

	We can extend the mapping $f$ to $\left( \mathcal{A} \cup \{ \intercal \}\right)^c$, by associating to the symbol $\intercal$ a cluster where none of the units are activated. If we consider the binary representation $\mu$ of $m$, erased symbols corresponds to erased clusters. An erased cluster is a cluster where all units are set to $0$. Formally the partial probe $\tilde{\mu}$ is such that for all $j\in [c]$ and $k \in [\ell ]$, \[ 
		\tilde{\mu}[(j-1)\ell+k] = \bigg \{ \begin{array}{cl} 0 & \mbox{if } j\in \{j_1, \dots , j_{c_e} \}\\
													\mu[(j-1)\ell+k] & \mbox{otherwise}	
	\end{array}\]

	\subsubsection{Hardness of the problem}

	An erasure is tantamount to lack of information, but associative memories can also retrieve messages from corrupt probes.

@@

	\subsubsection{Model}	
	We propose one model of message corruption. As all symbols in $\mathcal{A}$ all contains exactly $a$ ones, symbols which do not belong to $\mathcal{A}$ can be easily detected. We therefore restrict ourselves to random replacements by symbols in $\mathcal{A}$.

	A message $m$ can be recovered from a corrupt probe $\tilde{m}$ where some symbols, say $c_e$, of indexes $\{j_1, \dots , j_{c_e} \}$ have been replaced by random elements of $\mathcal{A}$. We have $\tilde{m} \in \mathcal{A}^c$.

@@

	\subsubsection{Hardness of the problem}

@@

	\subsubsection{Model}

@@

	We provide a formula for error rate in case of retrieving messages from a corrupt probe.

	%Messages can be corrupted, our models for errors is : given $c_e$ the number of erroneous clusters, we draw at random $c_e$ clusters and in each of these draw at random, uniformly, a new letter in the alphabet that will replace the old one (it could be the same).

	Knowing the density of the network, which follows formula (\ref{formula_density}), we can compute distributions for scores achieved by different type of units in different clusters.

	We can divide units in 5 groups with 3 characteristics. Subscript $g$ will indicate a unit being part of the message (it should be activated for the message to be retrieved), subscript $b$ the contrary (the unit should not be activated). Subscript $a$ means the unit is already activated, subscript $d$ that it is deactivated at the beginning. Subscript $c$ corresponds to a correct cluster, $e$ to an erroneous cluster. For example $s_{agc}$ designates the score achieved by an already \emph{a}ctivated unit (a) which should stay activated (g) in an uncorrupted cluster (c). Note that there is some redundancy.

	Correctly activated units in uncorrupted clusters achieve at least a score of $a(c-c_e - 1) + \gamma$. 
	Being connected to a given activated unit in a corrupted cluster occurs with probability $d$, since errors are identically distributed. Moreover corrects units see their scores shift, as do all activated units respectively thanks to being part of the original message and the memory effect $\gamma$. We can express those probability laws as : 
	For example an \textbf{a}ctivated \textbf{g}ood unit in a \textbf{c}orrect cluster starts with a base score of $ a(c - c_e - 1) + \gamma$. The probability it has to be connected to a random activated unit in an erroneous cluster is $d$. Thereby its law is :

@@

	Correct units in corrupted clusters achieve at least a score of $a(c - c_e)$. 
	$P(n_{dge} = a(c - c_e) + x) = {a (c_e - 1) \choose x} d^x (1-d)^{a (c_e - 1)-x}$

@@

	%$P(n_{be} = x) = P(n_{bc} = x) = {c - 1 \choose x} d^x (1-d)^{c- 1 -x}$
	Finally $P(n_{dbe} = x) = P(n_{dbc} = x) = {a(c - 1) \choose x} d^x (1-d)^{a(c- 1) -x}$

@@

	The probability for all good units of a correct cluster (resp. of an erroneous cluster) to achieve a score equal or greater than $x$, with at least one having a score of $x$ is $P_{ac} (x) = \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = x)^k P(n_{agc} > x)^{a-k}$ (resp. $P_e(x) = \sum_{k = 1}^a { a \choose k } P(s_{dge} = x)^k P(s_{dge} > x)^{a-k}$)

%	\begin{align*}
%	&P_{retrieve} = \left [ \sum_{n = 1}^{a(c-1) + \gamma} \left [ \sum_{k = 1}^{a} { a \choose k }   P(n_{agc} = n)^k P(n_{agc} > n)^{a-k} \right ] P(n_{bc} < n)^{l-a} \right ]^{c - c_e}\\ 
%	 &\times \left [ \sum_{n = 1}^{a(c-1) +\gamma} \left[ \sum_{k = 1}^a { a \choose k } P(n_{ge} = n)^k P(n_{ge} > n)^{a-k} \right ] P(n_{be} < n)^{l-2a} P(n_{abe} < n)^a \right ]^{c_e} \numberthis \label{from_err_th}
%	\end{align*}
		\begin{align*}
	&P_{retrieve} = \left [ \sum_{x = 1}^{a(c-1) + \gamma} P_{ac}(x) P(s_{dbc} < x)^{\ell-a} \right ]^{c - c_e}\\ 
	 &\times \left [ \sum_{x = 1}^{a(c-1) +\gamma} P_e(x) P(s_{dbe} < x)^{\ell-2a} P(s_{abe} < x)^a \right ]^{c_e} \numberthis \label{from_err_th}
	\end{align*}

@@

	As information on a message is shared across edges, networks with higher values of $a$ are more resilient to damages (as information on one message is carried by $a^2 {c \choose 2}$ edges. $v_c$ will indicate a correct unit that should be activated, $v$ the other type of units.
